---
title: "Shapes of Distributions"
output: learnr::tutorial
tutorial:
  id: "StatPREP/Nicole_lab_01"
  version: 0.1
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)

tutorial_options(exercise.checker = checkr::check_for_learnr)
# library(checkr)
# library(submittr)
library(dplyr)
library(ggformula)
library(openintro)
data(countyComplete)
Counties <- countyComplete
```

#<!--
#```{r child = system.file("shiny/login/login-mockup.Rmd", package = "submittr")}
#```
#-->

## Introduction

One of the Big Ideas of statistics we're studying in this course is called *descriptive statistics.*

The goal of descriptive statistics is to learn about and describe how things vary. What do we mean by "things"? It might be a bit more clear to say, "statistics describes how characteristics vary across the objects being studied."

In this lab, the "objects being studied" are the 3000 or so individual counties in the US. The "characteristics" are quantities tabulated for each county by the US Census Bureau, for instance the population of the county, or the average income of the people in the county.

You'll be using a data frame put together by the OpenIntro authors based on materials on the Census Bureau's web site. We'll be doing two things in this lab:

1. Getting familiar with the organization of the county data frame.
2. Looking at a few variables about counties and describing their *distributions*. 

## The `Counties` data frame 

The data frame is called `countyComplete`. Run the command in the following block to take a look at the data frame.


```{r get-data, exercise=TRUE}
countyComplete
```
 
Answer the following questions about the data frame.
```{r quiz1}
quiz(
  question("How many rows does the `countyComplete` data frame have?",
    answer("315"),
    answer("54"),
    answer("10"),
    answer("3143", correct = TRUE),
    answer("Cannot be determined.")
  ),
  question("How many variables does the `coutyComplete` data frame have?",
    answer("315"),
    answer("10"),
    answer("53", correct = TRUE),
    answer("3143"),
    answer("Cannot be determined.")
),
  
question("Scroll through the data to find the fraction of the population in Baldwin County, Alabama who completed high school.",
         answer("77.0%"),
         answer("87.6%", correct=TRUE),
         answer("85.3%"),
         answer("Cannot be deterimined."))
)
```



### Cases and variables

When you are first encountering a new data frame, you want to make sure you can answer two basic questions:

1. What kind of thing does each row correspond to? This is sometimes called a *case* or a *unit of observation*.
2. What does each variable describe? Sometimes this is pretty obvious, for instance with the variables called `state` and `name`. Other times you can make a reasonable guess, as with `pop2010`. 

Look back at the data frame on the previous page and try to make reasonable guesses about `age_under_5` and `veterans`.

```{r quiz2}
quiz(
  question("What do you think the variable `age_under_5` represents?",
    answer("The number of children in a county who are under 5 years old."),
    answer("The percentage of the county's population who are children under 5 years old.", correct = TRUE),
    answer("The average age of all children in the county who are under 5 years old."),
    answer("The number of children per family who are under 5 years old."),
    answer("Cannot be determined.")
  ),
  question("What do you think the variable `veterans` represents?",
    answer("The number of veterans living in the county", correct = TRUE),
    answer("The percentage of the county's population who are veterans"),
    answer("The number of people in the county who are receiving health care from the US Veterans' Administration"),
       answer("Cannot be determined.")
)
)
```


### Documentation
 
When data are properly collected and organized, you shouldn't have to make guesses about what is the unit of observation or exactly what each of the variables mean. Instead, the data should come with documentation that describe this precisely. 

The US Census Bureau is pretty careful to provide documentation. For the `countyComplete` data frame, you can access this easily with the R command, `help(countyComplete)`. Other sources of data provide documentation in other ways. 

Scan through the documentation of `countyComplete`.
```{r get_documentation, exercise = TRUE}
help(countyComplete)
```

### Assignment for your Lab 1 Report (Part 1)

Write answers to each of the following questions in your Lab 1 Report. Use complete sentences.

1. Sometimes data documentation here isn't very accurate or complete. Use the documentation you just obtained in the previous code block to look up the description of the `veterans` variable. Is it correct? Why or why not?

2. Use the data documentation and the data frame to look at the `mean_work_travel` information. (You can revisit the data frame using the code box below.) What do you think the *units* of time are (e.g., hours per year, minutes per week, etc.)? How or why did you choose these units?

```{r review data, exercise=TRUE}
countyComplete
```

## Some variables of interest

You're going to be looking at a few of the variables in the `countyComplete` data.

1. `foreign_born`
#. `persons_per_household`
#. `mean_work_travel`
#. `no_move_in_one_plus_year`

Look back at the documentation for `countyComplete` to make sure you understand what these variables stand for. You may want to write the information down on a piece of paper.

### Describing a variable

We want to describe each of these variables. Just as you might describe a person in many ways -- height, hair color, sex, personality, ... -- there are many ways to describe a variable.

For example, for any particular variable:
1. What's a small value? What's a big value?
2. What's a typical value?
3. How much variation is there?

A *statistic* is a number that can help to answer questions like these. There are many different statistics because there are many different questions to ask. Sometimes, there are multiple statistics that can be used to answer a given question in slightly different ways. 

In this lab, we'll look at two statistics used to describe a typical value: the *mean* and the *median*. 

We'll also look at another kind of description of a variable:

4. The *shape* of the distribution. There are many possible kinds of shape, but here we'll focus on one of the most basic description, whether the shape is *symmetric* or *skew*. 

### Foreign born

As you know, some people living in the US were born here. Others were born elsewhere. The `foreign_born` variable in the `countyComplete` data frame gives the percentage of the people living in each county who were born outside the US.

Before going on, think about what you expect about this variable. What's the percentage of foreign born in a typical county? What percentage would be pretty large when comparing counties to one another? What percentage would be pretty small?

Write down your answers on a scrap of paper. It doesn't matter whether you're right or wrong, just write down your best guess. The point here is that, when you look at the statistics, you can compare what they show to what you thought before you saw the actual data.

The following commands will calculate the mean and median of the `foreign_born` variable and show the shape.

```{r foreign-born, exercise = TRUE}
df_stats( ~ foreign_born, data = countyComplete, mean, median)
gf_histogram( ~ foreign_born, data = countyComplete)
```

This shape is called "skew". The word skew means that there are a small number of counties that have values that are far away from the values for the majority of counties. In particular, this is a "right-skew" distribution, which is to say that there are a handful of counties that have values for `foreign_born` that are two or three or more times larger than the values for most of the counties.

The mean and the median are statistics quantifying typical values for the variable. You can see from the computer output that the mean percentage of foreign born is about 4.4%. The median is 2.4%. The difference between the mean and median reflects the shape of the distribution. For a right-skew distribution, the mean is bigger than the median. Here, the mean is about twice the median.

### Moved recently?

The awkwardly named `no_move_in_one_plus_year` variable describes what fraction of the county's population did not move (change the location at which they live) at any point between 2006 and 2010. Among other things, a low value might mean that a large fraction of people live in unstable households.

```{r moved_recently, exercise = TRUE}
df_stats( ~ no_move_in_one_plus_year, data = countyComplete, mean, median)
gf_histogram( ~ no_move_in_one_plus_year, data = countyComplete)
```

The distribution of `no_move_in_one_plus_year` is skew. More specifically, it is *left-skew*, which means that some counties have values that are much lower than the majority of counties. Left-skew distribution have a mean that is *less* than the median; the opposite of the situation in right-skew distributions.

One way to think about much skewness there is the size of the difference between the mean and median. 

### Assignment for your Lab 1 Report (part 2)
Answer the following questions as if you are having a discussion with your boss, who has not taken a statistics course. Use complete sentences.
1. Describe what it means for a county to have a relatively low number in the `no_move_in_one_plus_year` category.
2. Describe what it means for a county to have a relatively high number in the `no_move_in_one_plus_year` category.

## Not skewed: symmetric

Many variables are *symmetric*, that is, neither left- nor right-skew.


```{r work-travel, exercise = TRUE}
df_stats( ~ mean_work_travel, data = countyComplete, mean, median)
gf_histogram( ~ mean_work_travel, data = countyComplete)
```

For a symmetric distribution, the mean and median are just about the same.

## Extremely skew

Some variables have very strongly skew distributions. For instance, the population of the counties varies tremendously, with a handful of counties having populations of one-million or more, while most counties have only about 20,000 people. 

```{r pop2010, exercise = TRUE}
df_stats( ~ pop2010, data = countyComplete, mean, median)
gf_histogram( ~ pop2010, data = countyComplete)
```

For such strongly skewed distributions, most of the counties are much closer to the median than that mean value. For this reason, it's appropriate to use a median to give a sense of a typical value.

It's especially common for variables that deal with money, population, or geographical area to be strongly right skew. This is true not just for US counties, but for countries of the world, or even for the income of individual people.

### Assignment for your Lab 1 Report (part 3)

Here are the names of the variables in the `countyComplete` data frame. 

```{r}
names(countyComplete)
```

Find examples of variables (other than the ones we've already seen) that have the designated shapes. Use the code block below to see the different distributions.
* a symmetric distribution
* a right-skew distribution

For each new example, write a short paragraph in which you:
* Identify the variable by name and describe its meaning.
* Describe the shape of the distribution for that variable.
* Were you surprised by the shape or did it make sense to you? Write a sentence or two explaining your answer.

If you find any that are left-skew, write those down too.

Run the following command as many times as you wish. You just have to replace the word VARIABLE with the actual name of a variable.

Tip from Nicole: You can highlight the name of a variable in the list above, then use CTRL-C to copy it and CTRL-V to paste it into the code below. Leave off the quotation marks!

```{r explore, exercise = TRUE}
df_stats( ~ VARIABLE, data = countyComplete, mean, median)
gf_histogram( ~ VARIABLE, data = countyComplete)
```

## To think about ...

For both skew and symmetrically shaped distributions, the median will be close to the large majority of values. Thus, the median is always a sensible description of a typical value. But for strongly skew distributions, the mean will be further than the median from the majority of values.



