---
title: "OpenIntro Lab 0"
output: learnr::tutorial
runtime: shiny_prerendered
tutorial:
  id: "StatPREP/OpenIntro/Lab0"
  version: 0.2
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
# DTK - I added this so that later chunks will have the dataset at the ready
Email_50 <- read.csv("https://nicolelang.github.io/Math1130/email50.csv")
library(checkr)
tutorial_options(exercise.checker = checkr::check_for_learnr)
library(submittr)
library(dplyr)
library(ggformula)
```

```{r child = system.file("shiny/login/login-mockup.Rmd", package = "submittr")}
```


## Labs?!? Do I need a white coat?
We usually think of a lab as a place where scientists (or science students) perform experiments. For students, the lab is a hands-on application of what is learned during the lecture part of their science course and a place where they can learn to use the tools used by scientists. 

In these labs, you'll be exploring and experimenting with the material learned in your textbook and/or classroom, and you'll also learn a bit about the tools used by statisticians. No, you don't need a white coat. You just need a browser, your notebook, and a pen or pencil.


## How is this going to happen?
This web site is connected (behind the scenes) to a server running software called R. (Yes, it's a single letter.) R can be used to do some pretty heavy-duty statistics stuff. The R software is free, and you can find a link to download both R and RStudio (an interface for using R) at <https://www.openintro.org/stat/labs.php>.

## Why aren't we using R and RStudio directly?
There are many reasons! One of them is that some of your classmates might not have their own computers, and students are not allowed to install software on NHCC's computers. Another is that R isn't a "point at the icon and click" kind of program. This web site can help with both of those reasons!

Are you ready? Let's explore!


## Step 1: Looking at data
Your textbook refers to several data sets in chapter 1. Let's find out how to look at these data sets.

### Finding the data 


We can have R pull in our data from a website. This means we don't have to spend a lot of time trying to type in hundreds of numbers! The box you see below contains the instructions for R to find the `email50` data from the web and put it into an R data set called "Email_50". (These instructions are called "R code")
The second line tells R to display the new R data set we created.
Click "Run Code" to see what happens!

```{r get-data, exercise=TRUE}
Email_50 <- read.csv("https://nicolelang.github.io/Math1130/email50.csv")
Email_50
```

The display you see above is called a *data frame*. 

Do you notice a triangular arrow in the upper right corner of the data table? What do you think that means? What about the information in the bottom row of the data display?

You can click on the triangle or the page numbers or "Previous" and "Next" to see what happens.

If you want to go back to the first version of the display, click "Start Over" in the upper left corner of the R Code box, then click "Run Code" again.

### Quick Quiz 1
```{r quiz1}
quiz(
  question("How many columns does the `Email_50` data frame have?",
    answer("5"),
    answer("21", correct = TRUE),
    answer("50"),
    answer("columns?")
  ),
  question("How many rows does the `Email_50` data frame have?",
    answer("5"),
    answer("21"),
    answer("50", correct = TRUE),
    answer("rows?")
),
  question("Each of the columns in the data frame represents a(n)...",
           answer("observation"),
           answer("experiment"),
           answer("email message"),
           answer("variable", correct=TRUE)
  ),
question("Each of the rows in the data frame represents a(n)...",
         answer("observation", correct=TRUE),
         answer("experiment"),
         answer("person"),
         answer("variable"))
)
```


## Narrowing it down

Many data sets come with more information than we can comfortably view on one screen. For now, we want to work with only one or two variables at a time.

This is how we tell R that we want just the `num_char` variable from the `Email_50` data frame. (Press "Run" to execute the command.) `num_char`  tells us the number of characters in each email message. 
Note: the values of this variable represent *thousands* of characters. So a value of 2.1 really means there were 2,100 characters in the email.


```{r restrict-to-num-char, exercise=TRUE}
select(Email_50, num_char)
```

Similarly, you can select two variables. For instance, here is the command to pull out the data used on p. 27 of your textbook. (Remember, click "Run Code" to execute the command.)

```{r new-table, exercise=TRUE}
select(Email_50, num_char, line_breaks)
```

Now your turn .... In the next box, type in the command to pull out `num_char`, `line_breaks`, and `dollar` from the Email_50 data frame. You can use "Run Code" to try out any command you compose. If you press "Submit," the tutorial will figure out if you've done the right thing and give you a hint if you haven't.

```{r new-table_2, exercise = TRUE}

```

```{r new-table_2-check}
#USER_CODE <- quote(select(Email_50, num_char, dollar, line_breaks))
#USER_CODE <- quote(select(Email_50, num_char))
code <- for_checkr(USER_CODE)
res <- if_empty_submission(code, message = "OK. You're starting with a blank. How about if you copy-and-paste the command from the previous code box into this one. That will give you a better starting point.")
res <- line_where(res, insist(F == select, "Use the `select()` function."))
res <- line_where(res, insist(is.data.frame(V), "Something's wrong."))
# Work around because Email_50 doesn't seem to be in the right scope.
dat <- arg_number(res, 1, insist(ncol(V) == 21, "The first argument to `select()` should be `Email_50`, the name of the data frame."))
if (failed(dat)) return(dat)

line_where(res, insist(ncol(V) == 3, "You should have all three variables in your command."),
           insist(all(names(V) %in% c("num_char", "dollar", "line_breaks")), 
                  "You have {{names(V)}} as variables. Are these what was asked for?"))

# line_where(res,
#            failif(ncol(V) == 2, "You should have the names of all three variables in the command."),
#            insist("dollar" %in% names(V), "You don't have the `dollar` variable in your command."),
#            insist("num_char" %in% names(V), "You don't have the `num_char` variable in your command."),
#            insist("line_breaks" %in% names(V), "You don't have the `line_breaks` variable in your command."),
#            insist(ncol(V) == 3, "You selected more than the three variables asked for."),
#            passif(TRUE, "Good!"))
```

```{r new-table_2-code-check}
3
```


## Looking ahead: graphical displays of data

The textbook has a scatterplot of `line_breaks` vs. `num_char`. Here's an R command to make that graph:
```{r canned_scatterplot, exercise = TRUE}
gf_point(Email_50, line_breaks ~ num_char)
```

This command is similar to the `select()` command you used before, but you can see two differences: 1) the function here is `gf_point()`, which is a function for making scatterplots. 2) Rather than the variable names being separated by a comma, they are put on either side of the squiggly "tilde" character ~. Tilde is the way you say "versus" in R.

Since we've already seen a scatterplot of `line_breaks` vs. `num_char`, let's look at something a bit different.

How do the number of characters in an email relate to the number of exclamation points in the message part of the email?

Here is a command to show the names of the columns in the `Email_50` data.


```{r column-names, exercise = TRUE}
names(Email_50)
```

Read through the names and try to figure out the name of the variable counting exclamation marks in the message. (There's also a variable counting exclamation marks in the email subject. That's not what we want.)

### Plotting...

We want to see if the number of exclamation points might depend on the number of characters in the email. Another way to say this is that we want to look at exclamation points *versus* number of characters. Generally, this means we want the number of exclamation points (the response variable) to be on the vertical or $y$-axis. We will put the number of characters (the explanatory variable) on the horizontal or $x$-axis.

In the R code below, replace `..explanatory_var..` and `..response_var..` with the appropriate variable names from the `Email_50` data. Then click "Run Code." When you have a graph that you think is right, press "Submit."

```{r plotting, exercise=TRUE}
gf_point(Email_50, ..response_var.. ~ ..explanatory_var..)
```

```{r plotting-check}
# USER_CODE <- quote(gf_point(Email_50, exclaim_mess ~ num_char))
# USER_CODE <- quote(gf_point(Email_50, exclaim_mess ~ num_char))
code <- for_checkr(USER_CODE)
res <- line_where(code, insist(TRUE)) # work around for `gf_point` not being in scope in chunk
# res <- line_where(code, insist(F == gf_point, "The `gf_point()` is the name of the scatterplot function. That's what you should use."))

# It seems that Email_50 is not always in the right environment for checkr to get a hold of it.
# Does this have to do with reading it in in the setup chunk. Bug in learnr?
# Work around? Instead of referring to Email_50, refer to properties of Email_50
# dat <- data_arg(res, insist(ncol(V) == 21, "You should be using the `Email_50` data set."))
# if (failed(dat)) return(dat)

# res <- formula_arg(res, message = "Make sure to use the squiggly tilde ~ to separate the names of the variables.")
# res <- check(res, 
#              insist("exclaim_mess" %in% all.vars(V), "The response variable should be `exclaim_mess`, not `{{f_lhs(V)}}`."),
#              insist("num_char" %in% all.vars(V), "The explanatory variable should be `num_char`, not `{{f_rhs(V)}}`."),
#              insist("exclaim_mess" == all.vars(V)[[1]], "Remember, the response variable goes to the left of the tilde, the response variable to the right."),
#              passif(TRUE, "That's it!")
#              )

return(res)
```


In the next lab, we will explore many different ways to look at data using graphs. We will also learn how to do some basic statistical calculations.



### Quick Quiz 2 (part 1)
```{r quiz2a}
quiz( question("Looking at the plot of *Number of Exclamation Points vs Number of Characters*, which of the following are true? (check all that apply)",
               answer("There is no association at all between these two variables."),
               answer("There is a moderate association between these two variables.", correct=TRUE),
               answer("There is a strong association between these two variables"),
               answer("If there is any association, it is negative."),
               answer("If there is any association, it is positive.", correct=TRUE),
               answer("There are no clumps of points on the graph."),
               answer("The points on the graph are mostly clumped together.", correct=TRUE)
      )
)
```

For part 2 of the quiz, click "Next Topic"

## Quick Quiz 2 (part 2)
For these questions, let's take another look at the graph of `line_breaks` vs. `num_char`.

Don't use your browser's "back" button! Create a fresh copy of the graph using the code here.

```{r canned_scatterplot2, exercise = TRUE}
gf_point(Email_50, line_breaks ~ num_char)
```



```{r quiz2b}
quiz(
  question("Look closely at the graph on page 27 of your textbook and compare it to the graph you just made here. What are the differences? (check all that apply)",
    answer("The axes are reversed."),
        answer("The axes labels are different.", correct=TRUE),
    answer("The data points are in different places."),
    answer("The points look different, but they're in the same places.", correct=TRUE),
    answer("The axes are different lengths."),
    answer("The axes are about the same lengths, but the tick marks are different.", correct=TRUE)
    
    ),

  
  question("Which of the following statements are true about the graph of *Number of Lines vs. Number of Characters*? (check all that apply)",
           answer("There is no association at all between these two variables."),
           answer("There is a moderate association between these two variables."),
           answer("There is a strong association between these two variables.", correct=TRUE),
           answer("If there is any association, it is negative."),  
           answer("If there is any association, it is positive.", correct=TRUE),
           answer("There are no clumps of points on the graph."),
           answer("The points on the graph are mostly clumped together.", correct=TRUE)
      )
)
```

### Instructions for submitting the Lab

At the top of each page in the tutorial there is a login ID and password box. You've been given your ID and password by the instructor. When you enter them into the boxes, the message will change to show that you are properly logged in. Any code that you "Submit" or any quizzes that you answer while logged in will be recorded automatically for the instructor. If you did the tutorial without logging in, press the "start over" label in the table of contents, login again, and repeat your work.
