---
title: "OpenIntro Lab 0"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Labs?!? Do I need a white coat?
We usually think of a lab as a place where scientists (or science students) perform experiments. For students, the lab is a hands-on application of what is learned during the lecture part of their science course and a place where they can learn to use the tools used by scientists. 

In these labs, you'll be exploring and experimenting with the material learned in your textbook and/or classroom, and you'll also learn a bit about the tools used by statisticians. No, you don't need a white coat. You just need a browser, your notebook, and a pen or pencil.

## How is this going to happen?
This web site is connected (behind the scenes) to a server running software called R. (Yes, it's a single letter.) R can be used to do some pretty heavy-duty statistics stuff. The R software is free, and you can find a link to download both R and RStudio (an interface for using R) at <https://www.openintro.org/stat/labs.php>.

## Why aren't we using R and RStudio directly?
There are many reasons! One of them is that some of your classmates might not have their own computers, and students are not allowed to install software on NHCC's computers. Another is that R isn't a "point at the icon and click" kind of program. This web site can help with both of those reasons!

Are you ready? Let's explore!


## Step 1: Looking at data
Your textbook refers to several data sets in chapter 1. Let's find out how to look at these data sets.

### Finding the data 


We can have R pull in our data from a website. This means we don't have to spend a lot of time trying to type in hundreds of numbers! The box you see below contains the instructions for R to find the `email50` data from the web and put it into an R data set called "Email_50". (These instructions are called "R code")
The second line tells R to display the new R data set we created.
Click "Run Code" to see what happens!

```{r get-data, exercise=TRUE}
Email_50 <- read.csv("https://nicolelang.github.io/Math1130/email50.csv")
Email_50

```

Do you notice a triangular arrow in the upper right corner of the data table? What do you think that means? What about the information in the bottom row of the data display?

You can click on the triangle or the page numbers ("Previous" and "Next") to see what happens.

If you want to go back to the first version of the display, click "Start Over" in the upper left corner of the Code box, then click "Run Code" again.

### Quick Quiz 1
```{r quiz1}
quiz(
  question("How many columns does the `Email_50` data set have?",
    answer("5"),
    answer("21", correct = TRUE),
    answer("50"),
    answer("columns?")
  ),
  question("How many rows does the `Email_50` data set have?",
    answer("5"),
    answer("21"),
    answer("50", correct = TRUE),
    answer("rows?")
),
  question("Each of the columns in the data table represents a(n)...",
           answer("observation"),
           answer("experiment"),
           answer("email message"),
           answer("variable", correct=TRUE)
  ),
question("Each of the rows in the data table represents a(n)...",
         answer("observation", correct=TRUE),
         answer("experiment"),
         answer("person"),
         answer("variable"))
)
```


## Narrowing it down

Many data sets come with more information than we can comfortably view on one screen. For now, we want to work with only one or two variables at a time.

This is how we tell R that we just want to look at the values for the `num_char` variable, which tells us the number of characters in each email message. 
Note: the values of this variable represent *thousands* of characters. So a value of 2.1 really means there were 2,100 characters in the email.

```{r reloading data} 
Email_50 <- read.csv("https://nicolelang.github.io/Math1130/email50.csv")
```

```{r restrict-to-num-char, exercise=TRUE}
Email_50$num_char
```

Do you like the way `num_char` displayed?
What if we want to have a nice tabular display like we did before, except with fewer columns? Let's see if we can look at both `num_char` and `line_breaks` (the data used on p.27 of your textbook).

```{r new-table, exercise=TRUE}
subset(Email_50, select=c("num_char","line_breaks"))
```

## Looking ahead: graphical displays of data
Since we've already seen a scatterplot of `line_breaks` vs. `num_char`, let's look at something a bit different.

How do the number of characters in an email relate to the number of exclamation points in the message part of the email?

Here are the names of the columns in the `Email_50` data.
```{r reloading2} 
Email_50 <- read.csv("https://nicolelang.github.io/Math1130/email50.csv")
```

```{r column-names}
names(Email_50)
```

Let's create the new data table. Replace `..variable..` in the R code below with the appropriate variable name for the number of exclamation points ("exclaim") in the *message*. Then click "Run Code."


```{r exlaim-table, exercise=TRUE}
subset(Email_50, select=c("num_char","..variable.."))
```

### Plotting...
We want to see if the number of exclamation points might depend on the number of characters in the email. Generally, this means we want the number of exclamation points (the response variable) to be on the vertical or $y$-axis. We will put the number of characters (the explanatory variable) on the horizontal or $x$-axis.

In the R code below, replace `..explanatory_var..` and `..response_var..` with the appropriate variable names from the `Email_50` data. Then click "Run Code."

```{r plotting, exercise=TRUE}
plot(x=Email_50$..explanatory_var.., y=Email_50$..response_var..)

```

In the next lab, we will explore many different ways to look at data using graphs. We will also learn how to do some basic statistical calculations.


### Quick Quiz 2
```{r quiz2a}
quiz( question("Looking at the plot of *Number of Exclamation Points vs Number of Characters*, which of the following are true? (check all that apply)",
               answer("There is no association at all between these two variables."),
               answer("There is a moderate association between these two variables.", correct=TRUE),
               answer("There is a strong association between these two variables"),
               answer("If there is any association, it is negative."),
               answer("If there is any association, it is positive.", correct=TRUE),
               answer("There are no clumps of points on the graph."),
               answer("The points on the graph are mostly clumped together.", correct=TRUE)
      )
)
```

#### Read this before answering the next question!
In the first part of Chapter 1, we discussed associations between variables. You can use the code box above to re-create the *Number of Lines vs. Number of Characters* graph from p.27 of your textbook.

You'll need to click "Start Over" in the upper left corner of the code box before you can create this plot.


```{r quiz2b}
quiz(
  question("Were you able to recreate the graph? (choose the best answer)",
    answer("Yes"),
        answer("Yes, but some details are different", correct = TRUE),
    answer("No.")
  ),
  question("Which of the following statements are true about the graph of *Number of Lines vs. Number of Characters*? (check all that apply)",
           answer("There is no association at all between these two variables."),
           answer("There is a moderate association between these two variables."),
           answer("There is a strong association between these two variables.", correct=TRUE),
           answer("If there is any association, it is negative."),  
           answer("If there is any association, it is positive.", correct=TRUE),
           answer("There are no clumps of points on the graph."),
           answer("The points on the graph are mostly clumped together.", correct=TRUE)
      )
)
```

### Instructions for submitting the Lab
I have no idea how this is going to work!