---
title: "OpenIntro Lab 0"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Labs?!? Do I need a white coat?
We usually think of a lab as a place where scientists (or science students) perform experiments. For students, the lab is a hands-on application of what is learned during the lecture part of their science course and a place where they can learn to use the tools used by scientists. 

In these labs, you'll be exploring and experimenting with the material learned in your textbook and/or classroom, and you'll also learn a bit about the tools used by statisticians. No, you don't need a white coat. You just need a browser, your notebook, and a pen or pencil.

## How is this going to happen?
This web site is connected (behind the scenes) to a server running software called R. (Yes, it's a single letter.) R can be used to do some pretty heavy-duty statistics stuff. The R software is free, and you can find a link to download both R and RStudio (an interface for using R) at <https://www.openintro.org/stat/labs.php>.

## Why aren't we using R and RStudio directly?
There are many reasons! One of them is that some of your classmates might not have their own computers, and students are not allowed to install software on NHCC's computers. Another is that R isn't a "point at the icon and click" kind of program. This web site can help with both of those reasons!

Are you ready? Let's explore!


## Step 1: Looking at data
Your textbook refers to several data sets in chapter 1. Let's find out how to look at these data sets.

### Finding the data 


We can have R pull in our data from a website. This means we don't have to spend a lot of time trying to type in hundreds of numbers! The box you see below contains the instructions for R to find the `email50` data from the web and put it into an R data set called "Email_50". (These instructions are called "R code")

```{r get-data, exercise=TRUE}
Email_50 <- ??????????????????

```

### Exercise with Code

*Here's an exercise with some prepopulated code as well as `exercise.lines = 5` to provide a bit more initial room to work.*

Now write a function that adds any two numbers and then call it:

```{r add-function, exercise=TRUE, exercise.lines = 5}
add <- function() {
  
}
```

## Topic 2

### Exercise with Hint

*Here's an exercise where the chunk is pre-evaulated via the `exercise.eval` option (so the user can see the default output we'd like them to customize). We also add a "hint" to the correct solution via the chunk immediate below labeled `print-limit-hint`.*

Modify the following code to limit the number of rows printed to 5:

```{r print-limit, exercise=TRUE, exercise.eval=TRUE}
mtcars
```

```{r print-limit-hint}
head(mtcars)
```

### Quiz

*You can include any number of single or multiple choice questions as a quiz. Use the `question` function to define a question and the `quiz` function for grouping multiple questions together.*

Some questions to verify that you understand the purposes of various base and recommended R packages:

```{r quiz}
quiz(
  question("Which package contains functions for installing other R packages?",
    answer("base"),
    answer("tools"),
    answer("utils", correct = TRUE),
    answer("codetools")
  ),
  question("Which of the R packages listed below are used to create plots?",
    answer("lattice", correct = TRUE),
    answer("tools"),
    answer("stats"),
    answer("grid", correct = TRUE)
  )
)
```

