---
title: "Mapping the counties"
output: learnr::tutorial
tutorial:
  id: "StatPREP/Nicole_lab_03"
  version: 0.1
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggformula)
library(openintro)
library(rlang)
library(maps)
library(mapproj)
data(countyComplete)
tutorial_options(exercise.checker = checkr::check_for_learnr)

countyMap <- function(var, 
                 projection = c("sinusoidal", "stereographic",
                 "azequalarea",
                 "azequidistant", "sinusoidal")) {
  projection <- match.arg(projection)
  var <- rlang::enexpr(var)
  varname <- rlang::expr_text(var)
  
  Shapes <- map_data('county') %>%
    rename(state = "region", county = "subregion")
  States <- map_data("state")

  Counties <- countyComplete %>% 
    mutate(var = !!var) %>%
    select(state, name, var) %>% 
    mutate(state = tolower(state), 
           county = gsub("( parish| county)$", "", tolower(name)),
           county = gsub("\\.", "", county))
  Both <- Shapes %>%  left_join(Counties)

  ggplot(Both, aes(x = long, y = lat, group = group, fill = var)) +
    geom_polygon() +
   scale_fill_continuous(guide = guide_legend(title = varname)) +
    geom_polygon(data = States, aes(x = long, y = lat, group = group), fill = NA, color = "gray", size = 0.05) +
    coord_map(projection = projection)
}
```

For the last two labs you've been looking at the county data without placing any of the counties on the map. 

The following chunk will generate a map of the counties. The color of the county shows the value of the variable. You can replace the `........` with any variable you like, e.g. `r paste(names(countyComplete)[-(1:3)], collapse = ", ")`


```{r showmap, exercise = TRUE, warning = FALSE, message=FALSE, out.width="100%"}
countyMap(.........)

```

### Your tasks

In performing these tasks, remember that you can put more than one `countyMay()` command in the command block, each on its own line. This will let you display multiple maps at the same time.

The variable `black` gives the proportion of the population in each county that is African American.

1. Make a map of the variable `black`.
2. Make a map of the quantity `black * pop2010`

Explain why the two maps show such different patterns.

3. Make a map of the absolute population growth from 2000 to 2010: that is `pop2010 - pop2000`
4. Make a map of the proportional population growth from 2000 to 2010, that is `(pop2010 - pop2000) / pop2000)`

Say why the maps are so different.

```{r include = FALSE, eval = FALSE}
library(checkr)
# uses the dbkeys.yml file to authenticate
library(submittr)
library(uuid)
options(tutorial.event_recorder =
          submittr::create_submit_to_db(table_name = "higgins",
                              password_table_name = "passwords"))
unique_session <- substr(UUIDgenerate(), 1, 8)
submittr::set_user(paste0("anonymous_", unique_session), "student", "Danny")
```



